Title here

### 2020-07-07 paper/paper.Rmd
make sure notation in the algorithm before matches with the ones in the global object

Now we formulate the problem in mathmetical notation. Projection pursuit finds the projection basis $A^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i$ is a $p \times 1$ vector, satisfying the following optimisation problem:

\begin{align}
I &= \arg \max_{A \in \mathcal{A}^{p \times d}} f(A) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in A: \lVert \mathbf{a}_i \rVert = 1 \}
\end{align}

via an iterative search: $\{A^1, A^2, \cdots \cdots A^J\}$. Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{A^1, A^1_1, A^1_2,\cdots, A^1_{k_1}, A^2, A^2_1, A^2_2, \cdots A^2_{k_2}, A^3, \cdots, A^{J-1}, A^{J-1}_{1}, \cdots A^{J-1}_{k_m} ,A^J  \}$$

@@

@posse1995projection used a stochastic direct search method, a random search algorithm to sample a candidate bases $A^{(j +1)}$ in the neighbourhood of the current basis $A^j$ defined by the radius of the p-dimensional sphere, $\alpha$. If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $A^{j + 1}$ along with other metadata and the algorithm stops. If no basis is found to have higher index value after the maximum number of tries $i_{max}$, the algorithm stops with nothing outputted. $h$ is the halfing parameter meaning if $h > 30$, the nieghbourhood parameter $\alpha$ will be reduced by half in the next iteration of the random search. The algorithm for a random search is summarised in Algorithm \ref{random-search}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $h = 0$, $loop = 1, i_{max}$\;
  \While{$i < i_{max}$}{
    $A^{(j + 1)} = A^{j} + \alpha A_{rand}$\;
    $I^{(j + 1)} = f(A^{(j+1)})$\;
    \eIf{$I^{(j+1)} > I^{j}$}{
      $A^{j+1} = A^{(j+1)}$\;
      }{
      $h = h + 1$\;
      }
    $i = i + 1$\;
  }
  \caption{random search: my first algorithm}
  \label{random-search}
\end{algorithm}

@@

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $loop = 1, i_{max}$\;
  \While{$i < i_{max}$}{
    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
    Find $A^{(J_1)}$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
    ...\;
    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
      \If{$p_{diff} > 0.001$}{
        $A^{j+1} = A^{(J_2)}$\;
        $I^{j+1} = I^{(J_2)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
      }
    $i = i + 1$\;
  }
  \caption{searchgeodesic}
  \label{search_geodesic}
\end{algorithm}

Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$. Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$. The simulated annealing algorithm works as follows. Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$ where $I^j$ and $I^{(j+1)}$ are the index value of the current and candidate basis respectively. The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $T_0$, $\alpha$\;
  \While{$i < i_{max}$}{
    $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
    $I^{(j+1)} = f(A^{(j+1)})$\;
    $T(i) = \frac{T_0}{\log(i + 1)}$\;
      \eIf{$I^{(j+1)} > I^{j}$}{
        $A^{j+1} = A^{(j+1)}$\;
        $I^{j+1} = I^{(j+1)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$,$i$ and relevant metadata}
      }{
        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$\;
        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
        \If{$\text{prob} > \text{rand}$}{
          \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
        }
      }
    $i = i + 1$\;
  }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

@@

*Iterators* indexes the data collected and it has a time series feature that prescribes the progression of the optimisation. Given the complexity of the algorithm, there could be multiple iterators. The simpliest iterator: *index per observation*, is a unique identifier for each observation in the data. For each level of iteration, we design two indices: *index per iteration* is fixed for each observations in one iteration and has an increment of one once a new iteration starts. *index within iteration* has an increment of one for each observation in an iteartion and starts over from one once a new iteration starts. Depend on the number of iteration level ($L$), the total number of iterators will differ and can be calculated as $1 + 2L$. In projection pursuit optimisation algorithms, there is only one level of iteartion and hence exists three iterators: $t_1$, $t_2$ and $t_3$. A sketch of the design of iterators is provided in Figure \ref{iterators}.

There could exist other parameters that are also of our interest but can't be classified as one of the three categories. They are defined under the fourth category: *other parameter of interest: $O$*. This could include variables that are specific for one particular algorithm. Below we show an example of defining the global object of projection pursuit guided tour.

Now we formulate the idea of global object in terms of mathmetical notation. Superscript $j$ is used to denote the number of iteration. Candidate basis are denoted using subscript on the right and interpolation basis are denoted using subscript on the left. The output basis doesn't have subscript. For example, $\mathbf{A}^1_{2}$ means the second candidate basis in the first iteration and ${}_{2}\mathbf{A}^1$ means the second interpolation basis that forms the tour path from the current basis $\mathbf{A}^0$ to the target basis $\mathbf{A}^1$. In the next iteration (iteration 2), $\mathbf{A}^1$ becomes the current basis, projection pursuit searches through $\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}$ to find the target basis $\mathbf{A}^2$ and the tour path from $\mathbf{A}^1$ to $\mathbf{A}^2$ is constructed using the interpolation bases ${}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2$. Below we arrange all the bases in a format that each coloumn represents an iteration in Equation \ref{eq:matrixA}.

When aggregated with information of index value and other interesting parameter, it is easier to write the basis as a column vector with its natual ordering as in Equation \ref{eq:vectorA}. Hence the full data structure can be written as in Equation \ref{eq:data-structure}

@@

@@

@@

\begin{equation}
\left[
\begin{array}{cccccccc}
\mathbf{A} & I & j & l & n & o_1 & \cdots &o_p\\
\hline
\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
\mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
\mathbf{A}^1_2 & I^1_2 & 1 & 2 & 3 & o_{1, 2} & \cdots & o_{p, 2}\\
\vdots &\vdots & \vdots &\vdots &\vdots &\vdots &\cdots &\vdots\\
\mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
\mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
{}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
{}_{2}\mathbf{A}^1 &{}_{2}I^1 & 1 & / & l_1 + 4&\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
{}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n^{*} &o_{1, n-1}& \cdots & o_{p, n-1}\\
\mathbf{A}^J & I^J & J & l_H + 1 & n^{**} & o_{1, n}& \cdots & o_{p, n}\\
\end{array}
\right]
\label{eq:data-structure}
\end{equation}

@@

An optimisation procedure can easily generate hundreds or thousands sampling points and the question becomes what are the points we want to explore... [the classification of searching or updating points goes here and people should be clear what part of the points goes to the plot - this is the data part not the plot part. ]

In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points

- *Searching points* include the observations recorded in the searching algorithm to find the target basis. The points for target bases is also included in the searching points and there is one such point per `tries`.

- *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value.

@@

In the projection pursuit guided tour, the decision variable is the projection matrix, the value of objective function is the index value. There are three iterators: `id` is the smallest ordering unit that increases by one for each observation; `tries` is updated once a search-and-interpolate step is finished. A third iterator `loop` is used to record the number of repetition in the search step and starts over from one at the beginning of a new `tries`. Three other parameters of our interest includes `method`, `alpha` and `info`.  `method` identifies the name of the searching method used and we are interested in comparing the performance between different algorithms under direct search. The neighbourhood parameter `alpha` controls the size of the sampling space and we are interested to understand how the searching space shrinks as the algorithm progresses. `info` labels different stages in the searching process. A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.

### Simulated data

>>>>

Now we formulate the problem in mathmetical notation. Projection pursuit finds the projection basis $\mathbf{A}^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i$ is a $p \times 1$ vector, satisfying the following optimisation problem:

\begin{align}
I &= \arg \max_{\mathbf{A} \in \mathcal{A}^{p \times d}} f(\mathbf{A}) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
\end{align}

via an iterative search: $\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$. Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{\mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, A^{J-1}, {}_{1}A^{J-1}, \cdots {}_{k_m}A^{J-1} ,A^J \}$$

Here we use left subscript for the index of the interpolation at each iteration i,e, ${}_{2}\mathbf{A}^1$ means the second interpolation basis in the first iteration. The right subcript is reserved for the index of searching points in the optimisation algorithm.

@@

@posse1995projection used a stochastic direct search method, a random search algorithm to sample a candidate bases $A^{j +1}_l$ in the neighbourhood of the current basis $A^j$ defined by the radius of the p-dimensional sphere, $\alpha$. If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $A^{j + 1}$ along with other metadata and the algorithm stops. If no basis is found to have higher index value after the maximum number of tries $l_{\max}$, the algorithm stops with nothing outputted. $c$ is the halfing parameter meaning if $c > 30$, the nieghbourhood parameter $\alpha$ will be reduced by half in the next iteration of the random search. The algorithm for a random search is summarised in Algorithm \ref{random-search}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $c = 0$, $l = 1$ $l_{\max}$, $\mathbf{A}^{j}$\;
  \While{$l < l_{\max}$}{
    $\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$\;
    $I^{j+1}_l = f(\mathbf{A}^{j+1}_l)$\;
    \eIf{$I^{j+1}_l > I^{j}$}{
      $\mathbf{A}^{j+1} = \mathbf{A}^{j+1}_{l}$\;
      }{
      $c = c + 1$\;
      }
    $l = l + 1$\;
  }
  \caption{random search: my first algorithm}
  \label{random-search}
\end{algorithm}

@@

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $l_{\max}$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
  \While{$l < l_{\max}$}{
    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
    Find $A^{J_1}_l$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
    ...\;
    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
      \If{$p_{diff} > 0.001$}{
        $A^{j+1} = A^{(J_2)}$\;
        $I^{j+1} = I^{(J_2)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
      }
    $i = i + 1$\;
  }
  \caption{searchgeodesic: this part has problem!}
  \label{search_geodesic}
\end{algorithm}

Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(l)$. Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{log(l + 1)}$. The simulated annealing algorithm works as follows. Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}$. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $\text{prob} = \\min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$ where $I^j$ and $I^{j+1}_l$ are the index value of the current and candidate basis respectively. The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $T_0$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
  \While{$l < l_{\max}$}{
    $\mathbf{A}^{j+1}_{l} = (1 - \alpha)\mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
    $I^{j+1}_l = f(\mathbf{A}^{j+1}_{l})$\;
    $T(l) = \frac{T_0}{\log(l + 1)}$\;
      \eIf{$I^{j+1}_l > I^{j}$}{
        $\mathbf{A}^{j+1} = \mathbf{A}^{j+1}_{l}$\;
        $I^{j+1} = I^{j+1}_l$\;
        \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$,$l$ and relevant metadata}
      }{
        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$\;
        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
        \If{$\text{prob} > \text{rand}$}{
          \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$, $l$ and relevant metadata}
        }
      }
    $l = l + 1$\;
  }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

@@

*Iterators* indexes the data collected and has a time series feature that prescribes the its natural ordering. Given the complexity of the algorithm, there could be multiple iterators. The simpliest iterator: *index per observation*, is a unique identifier for each observation in the data. For each level of iteration, we design two indices: *index per iteration* is fixed for each observations in one iteration and has an increment of one once a new iteration starts. *index within iteration* has an increment of one for each observation in an iteartion and starts over from one once a new iteration starts. A sketch of the difference between these three iterators is provided in Figure \ref{iterators}. Depend on the number of iteration level ($L$), the total number of iterators will differ and can be calculated as $1 + 2L$.

@@

In projection pursuit optimisation algorithms, there is one level of iteartion and hence exists three iterators: `id` indices each observation by a unique number; `tries` is the index per iteration iterator that gets updated once a search-and-interpolate step is finished; and `loop` is the index within iteration iterator and starts over from one at the beginning of a new `tries`.

There could exist other parameters that are also of our interest but can't be classified as one of the three categories. They are defined under the fourth category: *other parameter of interest: $O$*. In projection pursuit, three other parameters of our interest includes `method`, `alpha` and `info`.  `method` identifies the name of the searching method used and we are interested in comparing the performance between different algorithms under direct search. The neighbourhood parameter `alpha` controls the size of the sampling space and we are interested to understand how the searching space shrinks as the algorithm progresses. `info` labels different stages in the searching process. A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.

## Notation

Now we formulate the idea of global object using mathmetical notation. Let $\mathbf{A}^0$ denotes the starting basis. Superscript $j$ is used to denote the number of iteration (index per iteration). Candidate basis are denoted using subscript on the right and interpolation basis are denoted using subscript on the left.  For example, $\mathbf{A}^1_{2}$ means the second candidate basis in the first iteration and ${}_{2}\mathbf{A}^1$ means the second interpolation basis that forms the tour path from the current basis $\mathbf{A}^0$ to the target basis $\mathbf{A}^1$. In iteration two, $\mathbf{A}^1$ becomes the current basis and projection pursuit searches through $\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}$ to find the target basis $\mathbf{A}^2$. The tour path from $\mathbf{A}^1$ to $\mathbf{A}^2$ is constructed using the interpolation bases ${}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2$. All the bases together can be arranged in a format that each coloumn represents an iteration as in Equation \ref{eq:matrixA}.

@@

When aggregated with information of index value and other interesting parameter, it is easier to write the basis as a column vector as in Equation \ref{eq:vectorA} and the full data structure is presneted in Equation \ref{eq:data-structure}.

@@

\begin{equation}
\left[
\begin{array}{c|c|ccc|ccc}
\mathbf{A} & I & j & l & n & o_1 & \cdots &o_p\\
\hline
\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
\mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
\mathbf{A}^1_2 & I^1_2 & 1 & 2 & 3 & o_{1, 2} & \cdots & o_{p, 2}\\
\vdots &\vdots & \vdots &\vdots &\vdots &\vdots &\cdots &\vdots\\
\mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
\mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
{}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
{}_{2}\mathbf{A}^1 &{}_{2}I^1 & 1 & / & l_1 + 4&\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
{}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n^{*} &o_{1, n-1}& \cdots & o_{p, n-1}\\
\mathbf{A}^J & I^J & J & l_H + 1 & n^{**} & o_{1, n}& \cdots & o_{p, n}\\
\end{array}
\right]
\label{eq:data-structure}
\end{equation}

@@

<!-- An optimisation procedure can easily generate hundreds or thousands sampling points and the question becomes what are the points we want to explore... [the classification of searching or updating points goes here and people should be clear what part of the points goes to the plot - this is the data part not the plot part. ] -->

<!-- In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points -->

<!-- - *Searching points* include the observations recorded in the searching algorithm to find the target basis. The points for target bases is also included in the searching points and there is one such point per `tries`. -->

<!-- - *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value. -->

@@

## Simulated data
